### 2025-03-13 Reference Project/Draft EDA Airline Satisfaction_2025_03_04.Rmd
Add files via upload

---
title: "Draft EDA Airline Satisfaction"
output: html_document
date: "2025-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis

Importing Library

```{r}
library(dplyr)
library(ggplot2)
library(skimr)
library(DataExplorer)
```
Read Data

```{r}
data <- read.csv("airline_passenger_satisfaction.csv")
str(data)
```
```{r}
summary(data)
```

* ID : Unique passenger identifier
* Gender : Gender of the passenger (Female/Male)
* Age : Age of the passenger
* Customer Type : Type of airline customer (First-time/Returning)
* Type of Travel : Purpose of the flight (Business/Personal)
* Class : Travel class in the airplane for the passenger seat
* Flight Distance : Flight distance in miles
* Departure Delay : Flight departure delay in minutes
* Arrival Delay : Flight arrival delay in minutes
* Departure and Arrival Time Convenience : Satisfaction level with the convenience of the flight departure and arrival times from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* Ease of Online Booking : Satisfaction level with the online booking experience from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* Check-in Service : Satisfaction level with the check-in service from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* Online Boarding : Satisfaction level with the online boarding experience from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* Gate Location : Satisfaction level with the gate location in the airport from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* On-board Service : Satisfaction level with the on-boarding service in the airport from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* Seat Comfort : Satisfaction level with the comfort of the airplane seat from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* Leg Room Service : Satisfaction level with the leg room of the airplane seat from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* Cleanliness : Satisfaction level with the cleanliness of the airplane from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* Food and Drink : Satisfaction level with the food and drinks on the airplane from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* In-flight Service : Satisfaction level with the in-flight service from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* In-flight Wifi Service : Satisfaction level with the in-flight Wifi service from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* In-flight Entertainment	Satisfaction level with the in-flight entertainment from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* Baggage Handling : Satisfaction level with the baggage handling from the airline from 1 (lowest) to 5 (highest) - 0 means "not applicable"
* Satisfaction : Overall satisfaction level with the airline (Satisfied/Neutral or unsatisfied)

```{r}
any(is.na(data))
colMeans(is.na(data))
```
Null only exist for 0.3% and only in column Arrival Delay, therefore this would not significantly affecting the whole dataset. We'll fill the Arrival Delay null with mean and median of Arrival Delay whic is 0.

```{r}
data$Departure.Delay = as.numeric(data$Departure.Delay)
data$Arrival.Delay[is.na(data$Arrival.Delay)] <-0
any(is.na(data))
```
We succesfully eliminate null value in the dataset.

## Data Visualization

Now we visualize the distribution in each categorical column

```{r}
plot_bar(data,ncol=2,order_bar=T,ggtheme = theme_minimal())
```

We can notice that the gender distribution is equal. The type of travel is focusing on business with most of the customer are the returning customer rather than the first time. We can notice that focus in increasing satisfaction of loyal customer will bring significant impact to the overall of airline customer satisfaction rate.

Next we will visualize the distribution in each numerical column

```{r}
plot_histogram(data, ncol=2,ggtheme = theme_minimal())
```

We notice that there is some outliers in the Arrival Delay, Departure Delay, and the Flight Distance. Let's see in the boxplot format :

Flight Distance

```{r}
ggplot(data, aes(x=Flight.Distance)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

Departure Delay :

```{r}
ggplot(data, aes(x=Departure.Delay)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

Arrival Delay :

```{r}
ggplot(data, aes(x=Arrival.Delay)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

Median imputation is appropriate when the distribution of the data is skewed.

```{r}
data$Flight.Distance[data$Flight.Distance %in% boxplot(data$Flight.Distance)$out] = median(data$Flight.Distance)
data$Arrival.Delay[data$Arrival.Delay %in% boxplot(data$Arrival.Delay)$out] = median(data$Arrival.Delay)
data$Departure.Delay[data$Departure.Delay %in% boxplot(data$Departure.Delay)$out] = median(data$Departure.Delay)
```

## Feature Engineering

We notice that the age distribution have high variation, therefore we try to create the age_group column

```{r}
data = data %>% mutate(Age_Range = if_else(Age <= 16,"Children",
if_else(17 <= Age & Age <= 30,"Youth",
if_else(30 < Age & Age <= 45,"Middle Aged","Old"))))
```

```{r}
plot_bar(data$Age_Range,ncol=2,order_bar=T,ggtheme = theme_minimal())
```

Now we can interpret from the age group that the largest segment of our customer is belong to old age group which is >45 years old.

Another data we can group is the flight distance which currently right skewed

```{r}
data = data %>% mutate(Distance_Group = if_else(Flight.Distance < 1000,"Below 1000",
if_else(1000 <= Flight.Distance & Flight.Distance <= 2000,"1000-2000",
if_else(2000 < Flight.Distance & Flight.Distance < 3000,"2000-3000",
if_else(3000 < Flight.Distance & Flight.Distance < 4000,"3000-4000","Above 4000")))))
```

```{r}
plot_bar(data$Distance_Group,ncol=2,order_bar=T,ggtheme = theme_minimal())
```

## Exploratory Data Analysis

Our target is the satisfaction rate of the airlines services, therefore we need to identified how each variable are correlated to the satisfaction

```{r}
plot_bar(data,by='Satisfaction',ncol=2)
```

```{r}
data %>% group_by(Class, Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```

We notice that economy class passenger is having high customer base but with high percentage of dissatisfied customer rate. Therefore targeting this customer base for services improvement will highly affecting airline overall satisfaction rate.

```{r}
data %>% group_by(Type.of.Travel, Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```

Personal type of travel passenger is having a third of airline total customer with the number of satisfied customer is only 10% from it. Therefore it is also important customer segment to be improved.

```{r}
data %>% group_by(Customer.Type, Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```
76% of first time customer are feels dissatisfied with the airline services, this could be a problem because in the long term it will reducing airlines capability in increasing customer retention.

```{r}
data %>% filter(Customer.Type == 'First-time') %>% group_by(Class ,Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```

Most of the first-time customer feel dissatisfied in every class with notably higher in business and economy class.

```{r}
data %>% group_by(Distance_Group) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```

Based on the satisfaction rate in each of the distance group, we can notice that as the distance lower the rate of disatisfied customer is increasing. This become a problem because 60% of the airline customer is having journey below 1000.

Next we will try to analyze the service factors related to the customer satisfaction, starting with customer type :

```{r}

# Compute mean for each continuous column
filtered_data <- data %>% filter(Customer.Type == 'First-time') 
filtered_data <- subset(filtered_data, select = -c(ID, Flight.Distance, Age, Arrival.Delay, Departure.Delay))

mean_values <- sapply(filtered_data, function(x) if(is.numeric(x)) mean(x, na.rm = TRUE) else NA)
mean_values <- na.omit(mean_values)

mean_df <- data.frame(Variable = names(mean_values), Mean = mean_values)

# Remove NA values (non-numeric columns)
# Sort in descending order
mean_df <- mean_df[order(mean_df$Mean, decreasing = TRUE), ]

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```
We can identify that from the first time customer, the services that dissapointing them most in Departure and Arrival Time Convenience, Ease of Online Booking, and Wifi Service.

```{r}

# Compute mean for each continuous column
filtered_data <- data %>% filter(Type.of.Travel == 'Business') 
filtered_data <- subset(filtered_data, select = -c(ID, Flight.Distance, Age, Arrival.Delay, Departure.Delay))

mean_values <- sapply(filtered_data, function(x) if(is.numeric(x)) mean(x, na.rm = TRUE) else NA)
mean_values <- na.omit(mean_values)

mean_df <- data.frame(Variable = names(mean_values), Mean = mean_values)

# Remove NA values (non-numeric columns)
# Sort in descending order
mean_df <- mean_df[order(mean_df$Mean, decreasing = TRUE), ]

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```
We can identify that from the Personal type of travel customers, the services that dissapointing them most in Departure and Arrival Time Convenience, Ease of Online Booking, and Wifi Service.

```{r}

# Compute mean for each continuous column
filtered_data <- data %>% filter(Class == 'Economy') 
filtered_data <- subset(filtered_data, select = -c(ID, Flight.Distance, Age, Arrival.Delay, Departure.Delay))

mean_values <- sapply(filtered_data, function(x) if(is.numeric(x)) mean(x, na.rm = TRUE) else NA)
mean_values <- na.omit(mean_values)

mean_df <- data.frame(Variable = names(mean_values), Mean = mean_values)

# Remove NA values (non-numeric columns)
# Sort in descending order
mean_df <- mean_df[order(mean_df$Mean, decreasing = TRUE), ]

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```
We can identify that from the Economy class customers, the services that dissapointing them most in Online Boarding, Ease of Online Booking, and Wifi Service.

## Summary of EDA

- 43% of total passengers have liked the airline service. The majority of customers who are satisfied are from business class
- 57% of total passengers are not satisfied. The majority of them are from Economy class
- Only 18% of people have travelled for first time in our airline service. Out of those customers, nearly 76% of customers are not satisfied with our service which could create a chance in future of not returing of customers to our service again. When passengers are coming for first time, most of them have chosen the economy class
- Out of the customers who opted for business class, nearly 69% of customers are satisfied with the business class. Out of the customers who opted for economy class, nearly 81% of them are not satisfied with it. Customers who opted for economy plus class,nearly 75% of them are not satisfied with it.
- Whether people are coming for first time or not and also irrespective of their travel type, they are mostly not satisfied with the Economy and Economy Plus classes. There is strong need for the improvement of Economy class since 45% of total passengers are from economy class. Moreover, 60% of passengers travel less than 10k km and most of them prefer economy class.
- Passengers are not so satisfied with online booking, Online Boarding and the Wifi service mostly.
- 82% of total passengers are returning to the airline service. Out of those 105773 passengers,50% of passengers are from busines class and 42% of passengers are from economy class. Customer who are returning to our service are aged above 30 and make business travel and travel less than 1000 km mostly