### 2025-04-02 Draft Business Analytics Assignment Group 17.Rmd
Add files via upload
mohon direview bang

---
title: "Airline Customer Satisfaction - Group 17"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float: true
  pdf_document: default
date: "2025-03-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Background <a name="1"></a>

Lorem Ipsum

## Abstract <a name="1.1"></a>

**Framework for abstract:**

-   Introduction to airline industry

-   What the data is about

    -   customer satisfaction scores from 120,000+ passengers

    -   Nothing about the carrier

    -   Nothing about the price

-   What the research seeks to find

    -   Predict whether future customer would be satisfied given the details of service

    -   Determine the aspects of service that are important towards customer satisfaction

-   Methods of analysis

-   What the data tells us (findings)

**Draft of abstract:**

After industry-wide setbacks from the COVID-19 pandemic, the airline industry is experiencing a strong recovery in profitability. According to the International Air Transport Association (IATA), airlines worldwide have seen improved projections since 2024. The market is projected to showcase an annual growth rate (CAGR 2025-2029) of 4.36%, resulting in a projected market volume of US\$771.26bn by 2029, worldwide (Statista, 2024). However, low-cost carriers (LCCs) are shifting the traditional perceptions of value and service among travelers by offering lower fares with fewer amenities. As traditional full-service airlines compete in this market environment, understanding the main drivers of customer satisfaction becomes critical to maintain competitiveness.

In this research, we analyze data from over 120,000+ airline passengers, including information on passenger demographics, flight characteristics (delays and distance), as well as service evaluations. Since this data lacks information on the specific airlines and ticket pricing, we will focus on analyzing the service-related factors that universally impact passenger satisfaction. This research aims to:

1.  **Predict whether future customer would be satisfied given the details of service**
2.  **Determine the aspects of service that are important towards customer satisfaction**

In our analysis, we initially explore the dataset to understand the variables and their distributions, as well as detect missing values. Data preparation procedures such as cleaning, normalization, and transformation are conducted before analysis. We utilize Principal Component Analysis (PCA) to reduce dimensionality as well as to determine key variables. Clustering is then implemented to gain deeper insights on passenger groups and patterns of service satisfaction.**\<- placeholder aja, bisa diubah**

In summary.... **(hasil temuan)**

## Business Problem <a name="1.2"></a>

In this project, Group 17 serves as business consultants, with a focus on improving customer experience in the industry. Our responsibilities are from analyzing customer satisfaction statistics, developing measures to improve the airline's overall performance. Therefore, the business problem that we solve is "how can airlines improve its strategy of services and activities to increase customer satisfaction, especially between segments and main service areas, to achieve stronger competitive advantages?".

**The reason why we chose airline industry as our focus :**

-   Low-cost carriers (LCCs) are shifting the traditional perceptions of value and service among travelers by offering lower fares with fewer amenities.

-   Intense rivalry among airlines could be easily justified by customer satisfaction. Thus, higher customer satisfaction leads to increased client retention and overall business performance.

-   Concentrating on major market segments may provide a more general view of the aspects contributing to consumer satisfaction.

**Why focus on customer satisfaction?**

Enhance customer loyalty: improving on these key segments and service areas will enhance customer satisfaction that potentially increases customer loyalty rate.

Competitive Advantage: by addressing the service issue, the airline can stand out from the competitors.

## Setting Up the R Workplace <a name="1.3"></a>

Lorem IpsumBesides the Basic R Functions, this project uses multiple packages to enhance the R functionalities and improve the data analytics and visualization quality. This section explains all the packages and functions beyond the Basic R functionalities that we are using in this project in each step of the business analytics process from data cleaning to modeling.

* Data Cleaning and Preparation

**dplyr** : used to filter rows, select columns, sort data, create new columns, and summarize data.
**tidyr** : imported to be used to clean and tidy the data, such as reshaping the data from wide to long format.

* Data Exploration

**skimr** : designed to provide summary of statistical measures for each variables in data frames.
**DataExplorer** : used to display comprehensive and convenient interface for data exploration and data visualization.

* Data Transformation

* Data Visualization

**ggplot2** : a package that is useful to visualize data in a more complex and customizable way compared to the basic R plotting functions.
**corrplot** : used to visualize correlation matrix between two variables in a more visually appealing and informative way.

* Data Modeling

* Data Analysis

Below we will find the commands to install and import the required libraries:

```{r}
# Importing Library

library(dplyr)
library(ggplot2)
library(skimr)
library(DataExplorer)
library(corrplot)
library(tidyr)
```

## Dataset <a name="1.4"></a>

In this assignment, we use data set of [Airline Passenger Satisfaction Dataset](https://www.kaggle.com/datasets/mysarahmadbhat/airline-passenger-satisfaction) through Kaggle.The data consist of 24 variables from 129,880 passengers. Data cleaning and preparation are required to ensure the data is ready for analysis.

Firstly import all the data, csv file to be used in the data analysis process. The before using the data, it is important to undertand all data type included in the data set.Subsequently, understanding statistic perofile in the data set could help to prepare the data before using it.

```{r}
# Read Data
# please note that the csv needs to be put into the same folders as this file for ensuring everything works properly
# the dataset should be downloaded from the Kaggle, or unzip from our submission in BrightSpace.
data <- read.csv("airline_passenger_satisfaction.csv")

#Understanding the data types
str(data)
```

```{r}
# Summary of Data statistics
summary(data)
```

### Data Dictionary

Based on Kaggle, here is the definition of each column in the dataset. This is useful for later stage to understand the data characteristics and how to use it.

```{=html}
<style>
  table {
    width: 100%;
    border-collapse: collapse;
  }
  th {
    background-color: #87CEEB; /* Sky blue header */
    color: white;
    text-align: left;
    padding: 8px;
  }
  td {
    padding: 8px;
    text-align: left;
  }
  th, td {
    border: 1px solid #ddd;
  }
  th:first-child, td:first-child {
    width: 200px; /* Adjust field name column width */
  }
  th:last-child, td:last-child {
    width: 600px; /* Adjust description column width */
  }
</style>
```

| **Field Name** | **Description** |
|------------------------------------|------------------------------------|
| **ID** | Unique passenger identifier |
| **Gender** | Gender of the passenger (Female/Male) |
| **Age** | Age of the passenger |
| **Customer Type** | Type of airline customer (First-time/Returning) |
| **Type of Travel** | Purpose of the flight (Business/Personal) |
| **Class** | Travel class in the airplane for the passenger seat |
| **Flight Distance** | Flight distance in miles |
| **Departure Delay** | Flight departure delay in minutes |
| **Arrival Delay** | Flight arrival delay in minutes |
| **Departure and Arrival Time Convenience** | Satisfaction level with the convenience of the flight departure and arrival times from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Ease of Online Booking** | Satisfaction level with the online booking experience from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Check-in Service** | Satisfaction level with the check-in service from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Online Boarding** | Satisfaction level with the online boarding experience from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Gate Location** | Satisfaction level with the gate location in the airport from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **On-board Service** | Satisfaction level with the on-boarding service in the airport from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Seat Comfort** | Satisfaction level with the comfort of the airplane seat from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Leg Room Service** | Satisfaction level with the leg room of the airplane seat from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Cleanliness** | Satisfaction level with the cleanliness of the airplane from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Food and Drink** | Satisfaction level with the food and drinks on the airplane from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **In-flight Service** | Satisfaction level with the in-flight service from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **In-flight Wifi Service** | Satisfaction level with the in-flight Wifi service from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **In-flight Entertainment** | Satisfaction level with the in-flight entertainment from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Baggage Handling** | Satisfaction level with the baggage handling from the airline from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Satisfaction** | Overall satisfaction level with the airline (Satisfied/Neutral or unsatisfied) |

## Workflow: Machine-Learning Task Translation <a name="1.5"></a>

The business analysis focuses on what key parameters from passengers' perspective that define their satisfactory level towards airline services. The analysis will be conducted by perfroming predictive modelling (regression analysis) as well as classification. Run code belows to see the workflow of the analysis.

```{r}
file_path <- "Workflow_Group17.png"
knitr::include_graphics(file_path)
```

## Reflection and Limitation <a name="1.6"></a>

In this research, we aim to predict **whether future customer would be satisfied given the details of service** by determining **the aspects of service that are important towards customer satisfaction.** From the dataset, we hypothesize that service characteristics such as arrival and departure delays, on-board service, and comfort play a significant role towards customer satisfaction.

Given that the dataset does not provide information on the specific airline or ticket pricing, we will focus to analyzing the service-related factors that universally impact passenger satisfaction. Understanding and preparing the the dataset will be crucial, as it may contain missing values or quality issues. Hence, we will dedicate the next section on data exploration and preparation. In doing so, we will first visualize the variables to get an understanding of the distributions.

# Data Exploration and Preparation <a name="2"></a>

In this chapter, we will use visual and descriptive data exploration methods to gain a deeper understanding of the distributions and the characteristics of variables in our dataset. We will begin by cleaning the data from any null variables, as well has handle outliers in the data. Later in this chapter, we will continue to explore each variable and interpret features. Finally, we will conduct initial plotting of the data to show how the features are correlated to each other.

## Data Cleaning and Preparation <a name="2.1"></a>

We will start the exploration phase by checking for missing data using the `is.na()` function.

```{r check_missing_data}
any(is.na(data))
colMeans(is.na(data))
```

We find that null data only exists for 0.3% and only in the **Arrival Delay** column, therefore this would not significantly affecting the whole dataset. We'll fill the Arrival Delay null with the mean and median of Arrival Delay which is 0.

```{r eliminating_null_values}
data$Departure.Delay = as.numeric(data$Departure.Delay)
data$Arrival.Delay[is.na(data$Arrival.Delay)] <-0
any(is.na(data))
```

We succesfully replace the null values in the dataset. Let's check for numerical outliers using the `summary()` function, we will look for major differences between the median and the mean.

```{r summary_of_data}
# Extract summary statistics for numerical columns
numerical_summary <- data.frame(t(sapply(data[sapply(data, is.numeric)], summary)))

# Add a column to calculate the difference between mean and median
numerical_summary$Mean_Median_Difference <- abs(numerical_summary$Mean - numerical_summary$Median)

# Tabulate the output
knitr::kable(numerical_summary, caption = "Summary of Numerical Columns with Mean-Median Differences")

remove(numerical_summary)
```

We notice that there are large outliers in the Flight Distance, as well as in Arrival and Departure Delay. Let's see in the boxplot format :

**Flight Distance**

```{r}
ggplot(data, aes(x=Flight.Distance)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

**Departure Delay**

```{r}
ggplot(data, aes(x=Departure.Delay)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

**Arrival Delay**

```{r}
ggplot(data, aes(x=Arrival.Delay)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

From the boxplots above, we can see that Flight Distance is slightly skewed while Arrival Delay and Departure Delay is heavily skewed. We will use median imputation, which is appropriate when the distribution of the data is skewed.

```{r median_imputation}
# Imputation of flight.distance variable
data$Flight.Distance[which(
  data$Flight.Distance > (quantile(data$Flight.Distance, 0.75, na.rm = TRUE) + 1.5 * IQR(data$Flight.Distance, na.rm = TRUE))
)] <- median(data$Flight.Distance, na.rm = TRUE)

# Imputation of arrival.delay variable
data$Arrival.Delay[which(
  data$Arrival.Delay > (quantile(data$Arrival.Delay, 0.75, na.rm = TRUE) + 1.5 * IQR(data$Arrival.Delay, na.rm = TRUE))
)] <- median(data$Arrival.Delay, na.rm = TRUE)

# Imputation of departure.delay variable
data$Departure.Delay[which(
  data$Departure.Delay > (quantile(data$Departure.Delay, 0.75, na.rm = TRUE) + 1.5 * IQR(data$Departure.Delay, na.rm = TRUE))
)] <- median(data$Departure.Delay, na.rm = TRUE)
```

Now let's check the distribution of the data after the imputation.

```{r histogram_of_imputed_variables}
plot_histogram(data[ , c('Flight.Distance','Arrival.Delay','Departure.Delay')], ncol=2,ggtheme = theme_minimal())
```

Below, we will also take a look again at the boxplots for Departure Delay and Arrival Delay. Note that before the imputation, the plots were skewed to the point that the boxes were almost no longer visible.

```{r}
ggplot(data, aes(x=Departure.Delay)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

```{r}
ggplot(data, aes(x=Arrival.Delay)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

We can see that we have significantly normalized the skewness of the data. Thus, we have successfully eliminated all missing values and normalized all outliers in the data. Next, we will continue further into the data exploration phase.

## Data Exploration <a name="2.2"></a>

In this subchapter, we will begin by conducting descriptive univariate data analysis which is followed by association between numerical and categorical variables.

### Univariate Data Analysis

Now we visualize the distribution in each categorical column

```{r visualizing_categorical_variables}

# Bar chart will be used for categorical data
plot_bar(data, ncol=2, order_bar=TRUE, ggtheme = theme_minimal())
```

We can see that the gender distribution is equal. Furthermore, the majority type of travel is business which is also reflected in the customer type, where the majority is returning customers rather than first-time. Hence, increasing satisfaction of loyal customers will bring significant impact to the overall of airline customer satisfaction rate. Next, we will visualize the distribution in each numerical column. The ID column is left out as it is just a list of numbers.

```{r visualizing_numerical_variables}

# Numerical variables are visualized using histogram
plot_histogram(data[ , !names(data) %in% "ID"], ncol=2, ggtheme = theme_minimal())
```

From the visualizations above, it is not immediately clear which variables may be strongly influencing each other. Hence, we will continue the exploration by analyzing the associations between numerical variables.

### Association Between Two Numerical Variable

```{r corrplot, fig.width=14, fig.height=12}
numeric_data <- data[sapply(data, is.numeric)]
numeric_data <- na.omit(numeric_data)

# Compute correlation matrix
cor_matrix <- cor(numeric_data)

# Print the correlation matrix
print(cor_matrix)

# Visualize the correlation matrix
options(repr.plot.width=14, repr.plot.height=12)
corrplot(cor_matrix,
         method = "color",
         type = "upper",
         order = "hclust",
         tl.col = "black",
         tl.cex = 0.8,
         number.cex = 0.7,
         addCoef.col = "black",
         diag = FALSE)
```

Lorem Ipsum

```{r}
# Discover strong correlation
cor_matrix_flat <- as.data.frame(as.table(cor_matrix))
filtered <- subset(cor_matrix_flat, abs(Freq) > 0.5 & Var1 != Var2)
filtered <- filtered[!duplicated(t(apply(filtered[, 1:2], 1, sort))), ]

print(filtered)
```

In-flight comfort and service such as Seat Comfort, Leg Room Service, Cleanliness, Food and Drink, In-flight Service, In-flight Wifi Service, and In-flight Entertainment are highly correlated with each other. This is expected as they are all related to the comfort and service provided during the flight.

Digital service such as Ease of Online Booking, Online Boarding, and In-flight Wifi Service are also highly correlated with each other. This is expected as they are all related to the digital service provided by the airline.

```{r}
count_data <- as.data.frame(table(data$In.flight.Service, data$On.board.Service))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "In-flight Service vs On-board Service (in %)",
       x = "In-flight Service Rating",
       y = "On-board Service Rating",
       fill = "Percentage") +
  theme_minimal()
```

Lorem Ipsum

```{r}
count_data <- as.data.frame(table(data$In.flight.Entertainment, data$Food.and.Drink))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "In-flight Entertainment vs Food and Drink (in %)",
       x = "In-flight Entertainment Rating",
       y = "Food and Drink Rating",
       fill = "Percentage") +
  theme_minimal()
```

Lorem Ipsum

```{r}
count_data <- as.data.frame(table(data$In.flight.Entertainment, data$Cleanliness))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "In-flight Entertainment vs Cleanliness (in %)",
       x = "In-flight Entertainment Rating",
       y = "Cleanliness Rating",
       fill = "Percentage") +
  theme_minimal()
```

Lorem Ipsum

```{r}
count_data <- as.data.frame(table(data$In.flight.Entertainment, data$Seat.Comfort))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "In-flight Entertainment vs Seat Comfort (in %)",
       x = "In-flight Entertainment Rating",
       y = "Seat Comfort Rating",
       fill = "Percentage") +
  theme_minimal()
```

Lorem Ipsum

```{r}
count_data <- as.data.frame(table(data$Food.and.Drink, data$Cleanliness))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "Food and Drink vs Cleanliness (in %)",
       x = "Food and Drink Rating",
       y = "Cleanliness Rating",
       fill = "Percentage") +
  theme_minimal()
```

Lorem Ipsum

```{r}
count_data <- as.data.frame(table(data$Food.and.Drink, data$Seat.Comfort))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "Food and Drink vs Seat Comfort (in %)",
       x = "Food and Drink Rating",
       y = "Seat Comfort Rating",
       fill = "Percentage") +
  theme_minimal()
```

Lorem Ipsum

```{r}
count_data <- as.data.frame(table(data$Cleanliness, data$Seat.Comfort))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "Cleanliness vs Seat Comfort (in %)",
       x = "Cleanliness Rating",
       y = "Seat Comfort Rating",
       fill = "Percentage") +
  theme_minimal()
```

Lorem Ipsum

```{r}
count_data <- as.data.frame(table(data$In.flight.Wifi.Service, data$Ease.of.Online.Booking))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "In-Flight Wifi Service vs Ease of Online Booking (in %)",
       x = "In-Flight Wifi Service Rating",
       y = "Ease of Online Booking Rating",
       fill = "Percentage") +
  theme_minimal()
```

Lorem Ipsum

```{r}
ggplot(data, aes(x = Flight.Distance, y = Departure.Delay)) +
  geom_point(alpha = 0.2, color = "steelblue", size = 1.5) +
  labs(title = "Flight Distance vs Departure Delay",
       x = "Flight Distance",
       y = "Departure Delay") +
  geom_smooth(method = "lm", se = FALSE, color = "red", linewidth = 1) +
  theme_minimal()
```

Lorem Ipsum

```{r}
ggplot(data, aes(x = Flight.Distance, y = Arrival.Delay)) +
  geom_point(alpha = 0.2, color = "steelblue", size = 1.5) +
  labs(title = "Flight Distance vs Arrival Delay",
       x = "Flight Distance",
       y = "Arrival Delay") +
  geom_smooth(method = "lm", se = FALSE, color = "red", linewidth = 1) +
  theme_minimal()
```

### Association Between Numerical and Categorical Variable

Lorem ipsum

```{r}
data_subset <- data %>%
  select(Satisfaction, Age, Flight.Distance, Departure.Delay,
         Arrival.Delay)

# Convert to long format for faceting
data_long <- pivot_longer(data_subset, -Satisfaction, names_to = "Variable", values_to = "Value")

# Plot boxplots with facets
ggplot(data_long, aes(x = Satisfaction, y = Value, fill = Satisfaction)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_minimal() +
  labs(title = "Boxplots of Numeric Variables by Satisfaction",
       x = NULL, y = NULL) +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "none"
  )
```

Lorem Ipsum

```{r}
data_subset <- data %>%
  select(Satisfaction, Ease.of.Online.Booking, Check.in.Service, Online.Boarding,
         Gate.Location, On.board.Service, Leg.Room.Service, Flight.Distance)

# Convert to long format
data_long <- pivot_longer(data_subset, -Satisfaction, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Satisfaction)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Satisfaction",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

```{r}
data_subset <- data %>%
  select(Satisfaction, Seat.Comfort, In.flight.Service, Baggage.Handling,
         In.flight.Entertainment, In.flight.Wifi.Service, Food.and.Drink, Cleanliness)

# Convert to long format
data_long <- pivot_longer(data_subset, -Satisfaction, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Satisfaction)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Satisfaction",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

Lorem Ipsum

```{r}
data_subset <- data %>%
  select(Class, Age, Flight.Distance, Departure.Delay,
         Arrival.Delay)

# Convert to long format for faceting
data_long <- pivot_longer(data_subset, -Class, names_to = "Variable", values_to = "Value")

# Plot boxplots with facets
ggplot(data_long, aes(x = Class, y = Value, fill = Class)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_minimal() +
  labs(title = "Boxplots of Numeric Variables by Class",
       x = NULL, y = NULL) +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "none"
  )
```

Lorem Ipsum

```{r}
data_subset <- data %>%
  select(Class, Ease.of.Online.Booking, Check.in.Service, Online.Boarding,
         Gate.Location, On.board.Service, Leg.Room.Service, Flight.Distance)

# Convert to long format
data_long <- pivot_longer(data_subset, -Class, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Class)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Class",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

```{r}
data_subset <- data %>%
  select(Class, Seat.Comfort, In.flight.Service, Baggage.Handling,
         In.flight.Entertainment, In.flight.Wifi.Service, Food.and.Drink, Cleanliness)

# Convert to long format
data_long <- pivot_longer(data_subset, -Class, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Class)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Class",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

Lorem Ipsum

```{r}
data_subset <- data %>%
  select(Type.of.Travel, Age, Flight.Distance, Departure.Delay,
         Arrival.Delay)

# Convert to long format for faceting
data_long <- pivot_longer(data_subset, -Type.of.Travel, names_to = "Variable", values_to = "Value")

# Plot boxplots with facets
ggplot(data_long, aes(x = Type.of.Travel, y = Value, fill = Type.of.Travel)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_minimal() +
  labs(title = "Boxplots of Numeric Variables by Type of Travel",
       x = NULL, y = NULL) +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "none"
  )
```

Lorem Ipsum

```{r}
data_subset <- data %>%
  select(Type.of.Travel, Ease.of.Online.Booking, Check.in.Service, Online.Boarding,
         Gate.Location, On.board.Service, Leg.Room.Service, Flight.Distance)

# Convert to long format
data_long <- pivot_longer(data_subset, -Type.of.Travel, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Type.of.Travel)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Type of Travel",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

```{r}
data_subset <- data %>%
  select(Type.of.Travel, Seat.Comfort, In.flight.Service, Baggage.Handling,
         In.flight.Entertainment, In.flight.Wifi.Service, Food.and.Drink, Cleanliness)

# Convert to long format
data_long <- pivot_longer(data_subset, -Type.of.Travel, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Type.of.Travel)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Type of Travel",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

Lorem Ipsum

```{r}
data_subset <- data %>%
  select(Gender, Age, Flight.Distance, Departure.Delay,
         Arrival.Delay)

# Convert to long format for faceting
data_long <- pivot_longer(data_subset, -Gender, names_to = "Variable", values_to = "Value")

# Plot boxplots with facets
ggplot(data_long, aes(x = Gender, y = Value, fill = Gender)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_minimal() +
  labs(title = "Boxplots of Numeric Variables by Gender",
       x = NULL, y = NULL) +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "none"
  )
```

Lorem Ipsum

```{r}
data_subset <- data %>%
  select(Gender, Ease.of.Online.Booking, Check.in.Service, Online.Boarding,
         Gate.Location, On.board.Service, Leg.Room.Service, Flight.Distance)

# Convert to long format
data_long <- pivot_longer(data_subset, -Gender, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Gender)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Gender",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

```{r}
data_subset <- data %>%
  select(Gender, Seat.Comfort, In.flight.Service, Baggage.Handling,
         In.flight.Entertainment, In.flight.Wifi.Service, Food.and.Drink, Cleanliness)

# Convert to long format
data_long <- pivot_longer(data_subset, -Gender, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Gender)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Gender",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

Lorem Ipsum

```{r}
data_subset <- data %>%
  select(Customer.Type, Age, Flight.Distance, Departure.Delay,
         Arrival.Delay)

# Convert to long format for faceting
data_long <- pivot_longer(data_subset, -Customer.Type, names_to = "Variable", values_to = "Value")

# Plot boxplots with facets
ggplot(data_long, aes(x = Customer.Type, y = Value, fill = Customer.Type)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_minimal() +
  labs(title = "Boxplots of Numeric Variables by Customer Type",
       x = NULL, y = NULL) +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "none"
  )
```

Lorem Ipsum

```{r}
data_subset <- data %>%
  select(Customer.Type, Ease.of.Online.Booking, Check.in.Service, Online.Boarding,
         Gate.Location, On.board.Service, Leg.Room.Service, Flight.Distance)

# Convert to long format
data_long <- pivot_longer(data_subset, -Customer.Type, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Customer.Type)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Customer Type",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

```{r}
data_subset <- data %>%
  select(Customer.Type, Seat.Comfort, In.flight.Service, Baggage.Handling,
         In.flight.Entertainment, In.flight.Wifi.Service, Food.and.Drink, Cleanliness)

# Convert to long format
data_long <- pivot_longer(data_subset, -Customer.Type, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Customer.Type)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Customer Type",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

Lorem Ipsum

### Association Between Two Categorical Variable

```{r}
plot_bar(data,ncol=2)
```

Lorem ipsum

```{r}
plot_bar(data,by='Satisfaction',ncol=2)
```

Lorem ipsum

```{r}
plot_bar(data,by='Gender',ncol=2)
```

Lorem ipsum

```{r}
plot_bar(data,by='Customer.Type',ncol=2)
```

Lorem ipsum

```{r}
plot_bar(data,by='Type.of.Travel',ncol=2)
```

Lorem ipsum

```{r}
plot_bar(data,by='Class',ncol=2)
```

Lorem ipsum

## Feature Engineering <a name="2.3"></a>

We notice that the age distribution have high variation, therefore we try to create the age_group column

```{r}
data = data %>% mutate(Age_Range = if_else(Age <= 16,"Children",
if_else(17 <= Age & Age <= 30,"Youth",
if_else(30 < Age & Age <= 45,"Middle Aged","Old"))))
```

```{r}
plot_bar(data$Age_Range,ncol=2,order_bar=T,ggtheme = theme_minimal())
```

Now we can interpret from the age group that the largest segment of our customer is belong to old age group which is >45 years old.

Another data we can group is the flight distance which currently right skewed

```{r}
data = data %>% mutate(Distance_Group = if_else(Flight.Distance < 1000,"Below 1000",
if_else(1000 <= Flight.Distance & Flight.Distance <= 2000,"1000-2000",
if_else(2000 < Flight.Distance & Flight.Distance < 3000,"2000-3000",
if_else(3000 < Flight.Distance & Flight.Distance < 4000,"3000-4000","Above 4000")))))
```

```{r}
plot_bar(data$Distance_Group,ncol=2,order_bar=T,ggtheme = theme_minimal())
```

We can divide which customer experiencing delay or not based on departure delay and arrival delay

```{r}
data$delay.status = ifelse(data$Departure.Delay > 0 | data$Arrival.Delay > 0, "Delayed", "On Time")

plot_bar(data$delay.status,ncol=2,order_bar=T,ggtheme = theme_minimal())
```

We'll try to grouping the features into category of services that could be improved by the airline :

Flight General services : Food and Drink, In-flight Service
Flight Experience : In Flight Wifi Service, In Flight Entertainment
Digital services : Ease of Online Booking, Online Boarding
Comfort services : Seat Comfort, Leg Room Service, Cleanliness

```{r}
data$Flight.Service = rowSums(data[,c('Food.and.Drink','In.flight.Service')],na.rm = TRUE)
data$Flight.Entertainment = rowSums(data[,c('In.flight.Wifi.Service','In.flight.Entertainment')],na.rm = TRUE)
data$Digital.Service = rowSums(data[,c('Ease.of.Online.Booking','Online.Boarding')],na.rm = TRUE)
data$Comfort.Service = rowSums(data[,c('Seat.Comfort','Leg.Room.Service','Cleanliness')],na.rm = TRUE)

plot_histogram(data[ , c("Flight.Service", "Flight.Entertainment", "Digital.Service", "Comfort.Service")], ncol=2,ggtheme = theme_minimal())
```

Lorem Ipsum

## Data Tranformation <a name="2.4"></a>
Data transformation plays a key role to make sure all data is in the same scale and distribution. we will start  to take a look on our data distribution.

```{r}

library(ggplot2)
library(DataExplorer)

# Set default colors for histogram bars
update_geom_defaults("bar", list(fill = "steelblue", color = NA))

# Generate histogram with updated colors
plot_histogram(data, 
               ncol = 2, 
               ggtheme = theme_minimal() + theme(strip.background = element_blank(),
                                                 strip.text = element_text(color = "black")))

```
This part is to undertand the distribution of data before it is used. As we can see in the histogram, the data is not normally distributed. Therefore, we need to normalize the data.
There are several concerning variables in the dataset that are part of main focus in the data transformation process.
<ul>
<li>**arrival.delays** and **departure.delays** have highly-right skewed distribution where most values are concentrated to zero.</li>
<li>**flight.distance** has a right-skewed distribution concentrated between 500 and 1500, with a long tail up to 4000.</li>
<li>**age** has a right-skewed distribution concentrated between 20 and 50, with a long tail up to 80.But this variable is less concerning compared to the previous ones.</li>
</ul>
In the following step we will normalize the data by applying log transformation to the concerning variables.

```{r}
data$Departure.Delay.Duration = ifelse(data$Departure.Delay > 0, data$Departure.Delay, NA)
data$Arrival.Delay.Duration = ifelse(data$Arrival.Delay > 0, data$Arrival.Delay, NA)

data$Flight.Distance.log = log(1 + data$Flight.Distance)
data$Age.log = log(1 + data$Age)
data$Departure.Delay.Duration.log = log(1 + data$Departure.Delay.Duration)
data$Arrival.Delay.Duration.log = log(1 + data$Arrival.Delay.Duration)

plot_histogram(data[, c("Flight.Distance.log", "Age.log", "Departure.Delay.Duration.log", "Arrival.Delay.Duration.log")], ncol=2,ggtheme = theme_minimal())
```
Data transformation is performed with visualization task to see the distribution of the data after the transformation. This aims to normalize skewed distribution.
<ul>
  <li>**arrival.delays** and **departure.delays** are transformed into log scale to address the skewed issues.</li>
  <li>**flight.distance** and **age** are also transformed into log scale to address the skewed issues.</li>
</ul>
After log-transormation, normalize the distribution is then performed as follow.

```{r}
data$Flight.Distance.log.z = scale(data$Flight.Distance.log)
data$Age.log.z = scale(data$Age.log)
data$Departure.Delay.Duration.log.z = scale(data$Departure.Delay.Duration.log)
data$Arrival.Delay.Duration.log.z = scale(data$Arrival.Delay.Duration.log)

plot_histogram(data[, c("Flight.Distance.log.z", "Age.log.z", "Departure.Delay.Duration.log.z", "Arrival.Delay.Duration.log.z")], ncol=2,ggtheme = theme_minimal())
```

As we can see in the histogram, the data is now normalized with 0 as the center. This is important to ensure that the data is on the same scale and distribution. This helps the model learn the data better and make better predictions.

Here are the reasons why performing normalization is important before using the data:
<ol>
  <li><strong>Comparison Across Variables</strong><br>
      Normalization ensures the data is on the same scale. This helps in making the model comparable and supports easier further analysis.
  </li>
  <li><strong>Improved Model Performance</strong><br>
      Many algorithms perform better when input variables are standardized, as it prevents variables with larger ranges from dominating the analysis.
  </li>
  <li><strong>Normalization Beyond Log Transformation</strong><br>
      While log transformation reduces skewness and normalizes distributions, scaling ensures uniformity across variables, enabling fair use in statistical techniques.
  </li>
</ol>

## Initial Plotting <a name="2.5"></a>

Our target is the satisfaction rate of the airlines services, therefore we need to identify how each variable are correlated to the satisfaction. This is important to understand how each variable's behavior toward main objective of the analysis.
Following is the behavior of each variables towards satisfaction rate.

```{r}
plot_bar(data,by='Satisfaction',ncol=2)
```
According to the graphic above, we can see that:
<ul>
<li>**Gender**: Both male and female customers have a higher proportion of "Neutral or Dissatisfied" responses compared to "Satisfied" ones. This indicates there is a room of improvement to increase satisfied number</li>
<li>**Customer Type**: Returning customers exhibit a higher proportion of "Satisfied" responses than first-time customers, who seem mostly "Neutral or Dissatisfied". This might indicate that familiarity with the company positively influences satisfaction.</li>
<li> **Type of Travel**: Business travelers show a higher proportion of "Satisfied" responses compared to personal travelers.</li>
<li> **Class**: Business class has the highest proportion of "Satisfied" responses, whereas Economy and Economy Plus lean more towards "Neutral or Dissatisfied." This indicates higher satisfaction among premium customers. </li>
<li> **Age Range**: Older and middle-aged customers show higher satisfaction levels compared to youth and children, who lean towards dissatisfaction.</li>
<li> **Distance Group**:Longer trips seem to lead to higher satisfaction levels, which could indicate that services during these trips meet or exceed customer expectations. </li>
<li> **Delay Status**: the behavior of satisfaction between on-time and delayed status is not significantly different </li>
</ul>

```{r}
data_subset <- data %>%
  select(Satisfaction, Flight.Service, Flight.Entertainment, Digital.Service, Comfort.Service)

# Convert to long format
data_long <- pivot_longer(data_subset, -Satisfaction, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Satisfaction)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of New Numeric Variables by Satisfaction",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

Graphic above is telling the satisfaction distribution based on services category such as comfort, digital, entertainment, and flight services. We can see that the distribution of satisfaction is not significantly different between each category.

```{r}
data %>% group_by(Class, Satisfaction) %>% summarize(num=n()) %>% mutate(percentage=round(num*100/sum(num),2))
```
Based on the data above, **69.44%** of Business passengers showing satisfaction, while only *18.77%* of Economy passengers showing satisfaction and only **24.64%** of Economy Plus passengers are satisfied. This indicates that the airline services are more satisfying for business class passengers compared to economy class passengers. This data highlights the need for airlines to focus on enhancing the customer experience, particularly in Economy and Economy Plus, to achieve better satisfaction rates.

```{r}
data %>% group_by(Type.of.Travel, Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```

Personal type of travel passenger is having a third of airline total customer with the number of satisfied customer is only **10%** from it. This is much lower compared to the business type customers with **58.37%** satisfaction rate. Personal travel satisfaction levels indicate areas for improvement in service offerings or customer experience. Therefore it is also important customer segment to be improved trying to understand more satisfaction triggers from personal type of travelers.

```{r}
data %>% group_by(Customer.Type, Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))

```

**First-time Customers**: A majority, 76.03%, are neutral or dissatisfied, while only 23.97% report being satisfied.This could be a concern since it is indicating a risk from airline for not retaining first-time customers.<br>
<br>
**Returning Customers**: A majority, 58.37%, are satisfied, while 41.63% are neutral or dissatisfied. This indicates that returning customers are more satisfied with the airline services compared to first-time customers. This data highlights the need for airlines to focus on enhancing the customer experience, particularly for first-time customers, to achieve better satisfaction rates.

```{r}
data %>% filter(Customer.Type == 'First-time') %>% group_by(Class ,Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))

```
The table shows significant differences in first-timer passenger satisfaction across flight classes. Business Class shows the highest satisfaction, with nearly 40% satisfied, though more than half remain neutral or dissatisfied. Economy Class fares worse, with 85% of passengers neutral or dissatisfied, and Economy Plus performs the poorest, with 92% dissatisfaction. Enhancing services for Economy and Economy Plus could greatly improve overall customer satisfaction.<br>
<br>
In conclusion, mostly first-timer passengers feeling dissatisfied on each classes with notably higher in economy and economy plus class. This is a concern for the airline, as it indicates a need to improve services for first-time customers, to retain them and increase satisfaction rates.

```{r}
data %>% group_by(Distance_Group) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```
Previously in the barplot satisfaction, we can notice that  as the distance lower the rate of disatisfied customer is increasing. **Below 1000 distance is having the highest dissatisfaction rate**. From the data above, this condition could be a big concern since the majority of the airline customer is having journey below 1000 (more than 60%). This is a concern for the airline, as it indicates a need to improve services for short-distance travelers, to retain them and increase satisfaction rates.
<br>
Next we will try to analyze the service factors related to the customer satisfaction, starting with customer type :

```{r}

# Compute mean for each continuous column
filtered_data <- data %>% filter(Customer.Type == 'First-time') 
filtered_data <- subset(filtered_data, select = -c(ID, Flight.Distance, Age, Arrival.Delay, Departure.Delay, Arrival.Delay.Duration, Arrival.Delay.Duration.log,
                                                   Arrival.Delay.Duration.log.z, Departure.Delay.Duration, Departure.Delay.Duration.log, Departure.Delay.Duration.log.z,
                                                   Flight.Distance.log, Flight.Distance.log.z, Age.log, Age.log.z))

mean_values <- sapply(filtered_data, function(x) if(is.numeric(x)) mean(x, na.rm = TRUE) else NA)
mean_values <- na.omit(mean_values)

mean_df <- data.frame(Variable = names(mean_values), Mean = mean_values)

# Remove NA values (non-numeric columns)
# Sort in descending order
mean_df <- mean_df[order(mean_df$Mean, decreasing = TRUE), ]

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "#CFCFFF") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

We can identify that from the first time customer, the services that dissapointing them most in **Departure and Arrival Time Convenience, Ease of Online Booking, and Wifi Service**.While **Comfort Service** having the highest satisfaction value. <br>
This data can guide strategic improvements to enhance the first-time customer experience and increase satisfaction rates by focusing the improvement on those concerning satisfaction factors with the low score.

```{r}

# Compute mean for each continuous column
filtered_data <- data %>% filter(Type.of.Travel == 'Business') 
filtered_data <- subset(filtered_data, select = -c(ID, Flight.Distance, Age, Arrival.Delay, Departure.Delay, Arrival.Delay.Duration, Arrival.Delay.Duration.log,
                                                   Arrival.Delay.Duration.log.z, Departure.Delay.Duration, Departure.Delay.Duration.log, Departure.Delay.Duration.log.z,
                                                   Flight.Distance.log, Flight.Distance.log.z, Age.log, Age.log.z))

mean_values <- sapply(filtered_data, function(x) if(is.numeric(x)) mean(x, na.rm = TRUE) else NA)
mean_values <- na.omit(mean_values)

mean_df <- data.frame(Variable = names(mean_values), Mean = mean_values)

# Remove NA values (non-numeric columns)
# Sort in descending order
mean_df <- mean_df[order(mean_df$Mean, decreasing = TRUE), ]

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "#008080") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

We can identify that from the Personal type of travel customers, the services that dissapointing them most in **Departure and Arrival Time Convenience, In flight wifi service, Ease of Online Booking**.These areas could be prioritized for enhancement to improve overall satisfaction for business travelers. The top satisfaction factors is **Comfort service**.<br>
As you can see, personal type of travel customer satisfaction data is showing similar behaviour with the first-timer passengers. Thus, improving low-rated areas like WiFi availability and online booking ease may help elevate the overall customer experience for this segment.

```{r}

# Compute mean for each continuous column
filtered_data <- data %>% filter(Class == 'Economy') 
filtered_data <- subset(filtered_data, select = -c(ID, Flight.Distance, Age, Arrival.Delay, Departure.Delay, Arrival.Delay.Duration, Arrival.Delay.Duration.log,
                                                   Arrival.Delay.Duration.log.z, Departure.Delay.Duration, Departure.Delay.Duration.log, Departure.Delay.Duration.log.z,
                                                   Flight.Distance.log, Flight.Distance.log.z, Age.log, Age.log.z))

mean_values <- sapply(filtered_data, function(x) if(is.numeric(x)) mean(x, na.rm = TRUE) else NA)
mean_values <- na.omit(mean_values)

mean_df <- data.frame(Variable = names(mean_values), Mean = mean_values)

# Remove NA values (non-numeric columns)
# Sort in descending order
mean_df <- mean_df[order(mean_df$Mean, decreasing = TRUE), ]

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "#DC143C") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

The graph above is satisfaction data from Economy class customers. The biggest contributors to low satisfaction rate are **ease of online booking, wifi service, and online boarding**. While **Comfort Service** have the highest satisfaction rate among the Economy class customers. This data can guide strategic improvements to enhance the Economy class customer experience and increase satisfaction rates by focusing the improvement on those concerning satisfaction factors with the low score.

### Summary of Initial Plotting
<ol>
<li>43% of total passengers have liked the airline service. The majority of customers who are satisfied are from business class</li>
<li>57% of total passengers are not satisfied. The majority of them are from Economy class </li>
<li>Only 18% of people have travelled for first time in our airline service. Out of those customers, nearly 76% of customers are not satisfied with our service which could create a chance in future of not returing of customers to our service again. When passengers are coming for first time, most of them have chosen the economy class </li>
<li>Out of the customers who opted for business class, nearly 69% of customers are satisfied with the business class. Out of the customers who opted for economy class, nearly 81% of them are not satisfied with it. Customers who opted for economy plus class,nearly 75% of them are not satisfied with it. </li>
<li>Whether people are coming for first time or not and also irrespective of their travel type, they are mostly not satisfied with the Economy and Economy Plus classes. There is strong need for the improvement of Economy class since 45% of total passengers are from economy class. Moreover, 60% of passengers travel less than 10k km and most of them prefer economy class.</li>
<li>Passengers are not so satisfied with online booking, Online Boarding and the Wifi service mostly especially those who take economy class.</li>
<li>82% of total passengers are returning to the airline service. Out of those 105773 passengers,50% of passengers are from busines class and 42% of passengers are from economy class. Customer who are returning to our service are aged above 30 and make business travel and travel less than 1000 km mostly </li>
<li> Departure and arrival time covenience is considered big contributors to the customers dissatisfaction especially for first-timer passengers and personal travel type customers</li>
<li> Passengers mostly take short distance travel (below 1000 km) and it contributes to the high number of dissatisfaction. </li>
</ol>

## Reflections <a name="2.6"></a>

Lorem Ipsum

# Data Modeling <a name="3"></a>

Lorem Ipsum

## Modeling preparation <a name="3.1"></a>

Lorem Ipsum

## Classification Method <a name="3.2"></a>

Lorem Ipsum

## Logistic Regression <a name="3.3"></a>

Lorem Ipsum